---
title: "Storms in the North Atlantic"
format: 
  closeread-html:
    theme: cosmo
    css: ../../styles.css
    cr-style:
      narrative-background-color-sidebar: black
      section-background-color: white
      narrative-text-color-sidebar: white
      narrative-font-size: 2em
      #narrative-font-family: serif
---

```{r setup, include=FALSE}
library(tidyverse)

# number of storms per year (1980-2022)
storms_per_year = storms |>
  filter(year >= 1980) |> 
  mutate(id = paste0(name, "-", year)) |> 
  count(year, id, name = "count") |> 
  count(year, name = "count")
```


::::{.cr-section}


How many tropical storms per year have occurred in the North Atlantic between 1980 and 2022? @cr-plot1

:::{#cr-plot1}

```{r barchart1}
# version 1: start with an "empty" plot
storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = "#90909000") +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  scale_y_continuous(breaks = c(10, 20, 30), 
                     labels = c("10", "20", "30")) +
  theme_minimal() +
  theme(axis.text.y = element_text(color = "#00000000")) +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```

:::


This is the distribution of number of tropical storms per year in the 1980's  @cr-plot2

:::{#cr-plot2}
```{r barchart2}
# version 2: storms in 1980s
colors1 = c(rep("#136EED", 10), rep("#90909000", 33))

storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors1) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  scale_y_continuous(breaks = c(10, 20, 30), labels = c("10", "", "")) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


The number of tropical storms per year kept increasing in the 1990's. The year 1995 saw a record number of 21 storms. @cr-plot3

:::{#cr-plot3}
```{r barchart3}
# version 3: storms in 1990s
colors2 = c(rep("#bbbbbb", 10), rep("#136EED", 10), rep("#90909000", 23))

storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors2) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  scale_y_continuous(breaks = c(10, 20, 30), labels = c("10", "20", "")) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


The increasing trend continued during the 2000s, with a jaw dropping record of 29 storms in 2005. @cr-plot4

:::{#cr-plot4}
```{r barchart4}
# version 4: storms in 2000s
colors3 = c(rep("#bbbbbb", 20), rep("#136EED", 10), rep("#90909000", 13))

storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors3) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


In the 2010s decade the trend seemed to slow down a bit ...  @cr-plot5

:::{#cr-plot5}
```{r barchart5}
# version 5: storms in 2010s
colors4 = c(rep("#bbbbbb", 30), rep("#136EED", 10), rep("#90909000", 3))

storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors4) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


Only to reach a new all time high of 30 storms in 2020. @cr-plot6

:::{#cr-plot6}
```{r barchart6}
# version 6: storms in 2020-2022
colors5 = c(rep("#bbbbbb", 40), rep("#136EED", 3))

storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors5) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


So what was the overall trend between 1980 and 2022? @cr-plot7

:::{#cr-plot7}
```{r barchart7}
# version 7: all storms
storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = "#bbbbbb") +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


An increasing trend as you can tell. @cr-plot8

:::{#cr-plot8}
```{r barchart8, message=FALSE}
# version 8: all storms with fitted line
storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = "#bbbbbb") +
  geom_smooth(se = FALSE, linewidth = 1.4) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


::::
