---
title: "Storms in the North Atlantic"
format: 
  closeread-html:
    theme: cosmo
    css: ../../styles.css
    cr-style:
      narrative-background-color-sidebar: "#fcf7f0"
      section-background-color: white
      narrative-text-color-sidebar: black
      narrative-font-size: 2em
---

```{r setup, include=FALSE}
library(tidyverse)

# number of storms per year (1980-2022)
storms_per_year = storms |>
  filter(year >= 1980) |> 
  mutate(id = paste0(name, "-", year)) |> 
  count(year, id, name = "count") |> 
  count(year, name = "count")

# mean & median number of storms per decade
storms_per_decade = storms_per_year |>
  mutate(decade = as.numeric(paste0(str_sub(year, 1, 3), "0"))) |> 
  group_by(decade) |> 
  summarize(mean = mean(count),
            median = median(count),
            year = which.max(count),
            max = max(count)) |> 
  mutate(year = decade + year - 1)

# color palette
lowcolor = "#dccfb1" # low-light color
medcolor = "#BD835D" # highlight color per decade
maxcolor = "#ff9933" # color of maximum
```


::::{.cr-section}

What can the analysis of the __dplyr::storms__ data set reveal
about the number of tropical storms per year that occurred in the North Atlantic between 1980 and 2022? @cr-plot1

Let's take a look decade by decade.

:::{#cr-plot1}

```{r barchart1}
# version 1: start with an "empty" plot
storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = "#90909000") +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  scale_y_continuous(breaks = c(10, 20, 30), 
                     labels = c("10", "20", "30")) +
  theme_minimal() +
  theme(axis.text.y = element_text(color = "#00000000")) +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```

:::


In the 1980's the average number of tropical storms per year was 
`r storms_per_decade[1,2]`. @cr-plot2

:::{#cr-plot2}
```{r barchart2}
# version 2: storms in 1980s
colors1 = c(rep(medcolor, 10), rep("#90909000", 33))

storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors1) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  scale_y_continuous(breaks = c(10, 20, 30), labels = c("10", "", "")) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


The number of tropical storms per year kept increasing in the 1990's with an average of `r storms_per_decade[2,2]`. @cr-plot3

:::{#cr-plot3}
```{r barchart3}
# version 3: storms in 1990s
colors2 = c(rep(lowcolor, 10), rep(medcolor, 10), rep("#90909000", 23))

barchart3 = storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors2) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  scale_y_continuous(breaks = c(10, 20, 30), labels = c("10", "20", "")) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")

barchart3
```
:::


The year 1995 saw a record number of `r storms_per_decade[2,"max"]` storms. @cr-plot3max

:::{#cr-plot3max}
```{r barchart3max}
barchart3 +
  geom_col(data = filter(storms_per_decade, decade == 1990), 
           aes(x = year, y = max), fill = maxcolor)
```
:::



The increasing trend continued during the 2000's, averaging `r storms_per_decade[3,2]` storms per year @cr-plot4

:::{#cr-plot4}
```{r barchart4}
# version 4: storms in 2000s
colors3 = c(rep(lowcolor, 20), rep(medcolor, 10), rep("#90909000", 13))

barchart4 = storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors3) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")

barchart4
```
:::


With a jaw dropping record of `r storms_per_decade[3,"max"]` storms in 2005! @cr-plot4max

:::{#cr-plot4max}
```{r barchart4max}
barchart4 +
  geom_col(data = filter(storms_per_decade, decade == 2000), 
           aes(x = year, y = max), fill = maxcolor) 
```
:::


In the 2010's the trend seemed to slow down a bit with an average number of storms per year of `r storms_per_decade[4,2]` ...  @cr-plot5

:::{#cr-plot5}
```{r barchart5}
# version 5: storms in 2010s
colors4 = c(rep(lowcolor, 30), rep(medcolor, 10), rep("#90909000", 3))

storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors4) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


... only to go up again @cr-plot6

:::{#cr-plot6}
```{r barchart6}
# version 6: storms in 2020-2022
colors5 = c(rep(lowcolor, 40), rep(medcolor, 3))

barchart6 = storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = colors5) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")

barchart6
```
:::


reaching a new all time high of `r storms_per_decade[5,"max"]` storms in 2020. @cr-plot6max

:::{#cr-plot6max}
```{r barchart6max}
barchart6 +
  geom_col(data = filter(storms_per_decade, decade == 2020), 
           aes(x = year, y = max), fill = maxcolor)
```
:::


So what was the overall trend between 1980 and 2022? @cr-plot7

:::{#cr-plot7}
```{r barchart7}
# version 7: all storms
storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = lowcolor) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


An increasing trend as you can tell. @cr-plot8

:::{#cr-plot8}
```{r barchart8, message=FALSE}
# version 8: all storms with fitted line
storms_per_year |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = lowcolor) +
  geom_smooth(se = FALSE, linewidth = 1.4, color = medcolor) +
  scale_x_continuous(breaks = seq(1980, 2020, by = 5)) +
  theme_minimal() +
  labs(title = "Number of Storms per Year in the North Atlantic",
       x = "")
```
:::


::::
