---
title: "Mapping Storms in the North Atlantic"
format: 
  closeread-html:
    theme: cosmo
    css: ../../styles.css
    cr-style:
      narrative-background-color-sidebar: "#fcf7f0"
      section-background-color: white
      narrative-text-color-sidebar: black
      narrative-font-size: 2em
---

```{r setup, include=FALSE}
library(tidyverse)     # data wrangling and graphics
library(lubridate)     # for working with dates
library(sf)            # simple features
library(rnaturalearth) # maps data

storms2010 = storms |>
  filter(year == 2010) |> 
  mutate(date = as.Date(paste(year,month,day,sep="-"))) |> 
  select(name, date, month, lat, long, wind)

storms2010_aux = storms2010 |> 
  group_by(name) |> 
  summarize(date = first(date),
            first = str_sub(first(date), 6, 10),
            last = str_sub(last(date), 6, 10),
            lat = last(lat),
            long = last(long),
            month = month(date)) |> 
  arrange(first)

storms2010_per_month = storms2010_aux |> 
  count(month, name = "count")

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

n = length(unique(storms2010_aux$name))

storms2010_aux$color = gg_color_hue(n)

world_map = ne_countries(returnclass = "sf")
```



::::{.cr-section}

Tropical Storms in the North Atlantic, 2010 season. @cr-map1

:::{#cr-map1}
```{r map1}
ggplot() + 
  geom_sf(data = world_map) +
  coord_sf(xlim = c(-110,0), ylim = c(10, 60)) +
  geom_point(data = storms2010, 
             aes(x = long, y = lat), color = "#99999900") +
  labs(title = "Storms starting in June") +
  theme(title = element_text(color = "#99999900"),
        axis.title.x = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        panel.background = element_blank())
```
:::


The season started in June with `r storms2010_per_month$count[1]` storm: `r storms2010_aux$name[1]` @cr-storms1

:::{#cr-storms1}
```{r storms1}
name_aux = storms2010_aux$name[1]

ggplot() + 
  geom_sf(data = world_map) +
  coord_sf(xlim = c(-110,0), ylim = c(10, 60)) +
  geom_point(data = filter(storms2010, name == name_aux), 
             aes(x = long, y = lat), 
             color = storms2010_aux$color[1]) +
  geom_path(data = filter(storms2010, name == name_aux), 
             aes(x = long, y = lat), 
             color = storms2010_aux$color[1]) +
  geom_text(data = filter(storms2010_aux, name %in% name_aux),
            vjust = -0.5,
            aes(x = long, y = lat, label = name, color = name)) +
  labs(title = "Storms starting in June") +
  theme(panel.background = element_blank(),
        legend.position = "none")
```
:::


The month of July saw the starting of `r storms2010_per_month$count[2]` storms @cr-storms2

:::{#cr-storms2}
```{r storms2}
name_aux = storms2010_aux |> 
  filter(month == 7) |> 
  pull(name)

color_aux = storms2010_aux |> 
  filter(name %in% name_aux) |> 
  pull(color)

ggplot() + 
  geom_sf(data = world_map) +
  coord_sf(xlim = c(-110,0), ylim = c(10, 60)) +
  geom_point(data = filter(storms2010, name %in% name_aux), 
             aes(x = long, y = lat, color = name)) +
  geom_path(data = filter(storms2010, name %in% name_aux), 
             aes(x = long, y = lat, color = name)) +
  geom_text(data = filter(storms2010_aux, name %in% name_aux),
            vjust = -0.5,
            aes(x = long, y = lat, label = name, color = name)) +
  scale_color_manual(values = color_aux) +
  labs(title = "Storms starting in July") +
  theme(panel.background = element_blank(),
        legend.position = "none")
```
:::


The month of August saw the starting of `r storms2010_per_month$count[3]` storms @cr-storms3

:::{#cr-storms3}
```{r storms3}
name_aux = storms2010_aux |> 
  filter(month == 8) |> 
  pull(name)

color_aux = storms2010_aux |> 
  filter(name %in% name_aux) |> 
  pull(color)

ggplot() + 
  geom_sf(data = world_map) +
  coord_sf(xlim = c(-110,0), ylim = c(10, 60)) +
  geom_point(data = filter(storms2010, name %in% name_aux), 
             aes(x = long, y = lat, color = name)) +
  geom_path(data = filter(storms2010, name %in% name_aux), 
             aes(x = long, y = lat, color = name)) +
  geom_text(data = filter(storms2010_aux, name %in% name_aux),
            vjust = -0.5,
            aes(x = long, y = lat, label = name, color = name)) +
  scale_color_manual(values = color_aux) +
  labs(title = "Storms starting in August ") +
  theme(panel.background = element_blank(),
        legend.position = "none")
```
:::


The month of September saw the starting of `r storms2010_per_month$count[4]` storms @cr-storms4

:::{#cr-storms4}
```{r storms4}
name_aux = storms2010_aux |> 
  filter(month == 9) |> 
  pull(name)

color_aux = storms2010_aux |> 
  filter(name %in% name_aux) |> 
  pull(color)

ggplot() + 
  geom_sf(data = world_map) +
  coord_sf(xlim = c(-110,0), ylim = c(10, 60)) +
  geom_point(data = filter(storms2010, name %in% name_aux), 
             aes(x = long, y = lat, color = name)) +
  geom_path(data = filter(storms2010, name %in% name_aux), 
             aes(x = long, y = lat, color = name)) +
  geom_text(data = filter(storms2010_aux, name %in% name_aux),
            vjust = -0.5,
            aes(x = long, y = lat, label = name, color = name)) +
  scale_color_manual(values = color_aux) +
  labs(title = "Storms starting in September") +
  theme(panel.background = element_blank(),
        legend.position = "none")
```
:::


The month of October saw the starting of `r storms2010_per_month$count[5]` storms @cr-storms5

:::{#cr-storms5}
```{r storms5}
name_aux = storms2010_aux |> 
  filter(month == 10) |> 
  pull(name)

color_aux = storms2010_aux |> 
  filter(name %in% name_aux) |> 
  pull(color)

ggplot() + 
  geom_sf(data = world_map) +
  coord_sf(xlim = c(-110,0), ylim = c(10, 60)) +
  geom_point(data = filter(storms2010, name %in% name_aux), 
             aes(x = long, y = lat, color = name)) +
  geom_path(data = filter(storms2010, name %in% name_aux), 
             aes(x = long, y = lat, color = name)) +
  geom_text(data = filter(storms2010_aux, name %in% name_aux),
            vjust = -0.5,
            aes(x = long, y = lat, label = name, color = name)) +
  scale_color_manual(values = color_aux) +
  labs(title = "Storms starting in October") +
  theme(panel.background = element_blank(),
        legend.position = "none") 
```
:::



The 2010 season saw a total of `r n` storms. @cr-storms

:::{#cr-storms}
```{r storms}
ggplot() + 
  geom_sf(data = world_map) +
  coord_sf(xlim = c(-110,0), ylim = c(10, 60)) +
  geom_point(data = storms2010, 
             aes(x = long, y = lat, color = name)) +
  geom_path(data = storms2010, 
             aes(x = long, y = lat, color = name)) +
  labs(title = "Storms starting in 2010") +
  theme(panel.background = element_blank(),
        legend.position = "none")
```
:::


::::
